---
title: "Week 5:  Matrix Algebra"
author: "Stat 431"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("Canvas_Pages/Scripts/window_dressing.R"))
quiz_url <- "https://canvas.calpoly.edu/courses/16654/quizzes/14550"
library(tidyverse)
library(flair)
options(digits = 2)
```

```{css, echo = FALSE}
img {
  vertical-align:middle
}

body {
  text-shadow:none;
}


```




```{r, results = "asis", echo = FALSE}
make_summary_table(vid_time = "30 min", 
                   reading_time = "20 min", 
                   work_time = "40 min", 
                   canvas_activities = 1)
```


## Basics of Matrices

### Notation and Vocab

A **scalar** is a single number by itself.

In statistics, a scalar will usually be:

* A single data observation
* A parameter
* A summary statistic

In R, a scalar is just a single number; or equivalently, an "atomic vector":

```{r}
a <- 5
a
```

A **vector** is a set of scalars put together.

In statistics, a vector might be a set of *samples* from a single variable or a set of *observations of many variables* from a single sample.

In R, we use vectors all the time:

```{r}
vec <- c(1, 2, 3)
vec

head(iris$Sepal.Length)
```

A **matrix** is a two dimensional set of scalars; or equivalently, many vectors put together.

In statistics, a matrix usually represents observations from *one or more variables* (columns) for *many samples* (rows).

The **dimension of a matrix** ($m \times n$) is the number of rows ($m$) by the number of columns ($n$).
The elements of the matrix are often written as $a_{ij}$, as in
$$ {\bf A} \, = \, \left( \matrix{ a_{11} & a_{12} \\ a_{21} & a_{22} } \right) $$

You will sometimes hear of a $1 \times n$ matrix is called a *row vector* and an $m \times 1$ matrix is called a *column vector*.

Careful - in R, a *matrix* and a *vector* are different object types and sometimes behave differently!

```{r}
col_vec <- matrix(1:3, nrow = 3, ncol = 1)
row_vec <- matrix(1:3, nrow = 1, ncol = 3)

vec
col_vec
row_vec

class(vec)
class(col_vec)
class(row_vec)
```


A *square* matrix has the same number of rows as columns. A *diagonal* matrix has all zeros except on the diagonal.  One special *square, diagonal* matrix is the **identity matrix**:

$${\bf I_n} = \left( \matrix{1 & 0 & \ldots \\ 0 & 1 & \ldots \\ \vdots & \vdots & \vdots \\ \ldots & 0 & 1} \right)$$

```{r}
I <- diag(3)
I
```


### Basic math operations

For purposes, you'll need to understand the basics of doing math with matrices.
Here are a few quick tutorial (or refresher) options:


```{r, results = "asis", echo = FALSE}
req_vid("Matrix Math", url = 'https://www.youtube.com/watch?v=kqWCwwyeE6k#t=1m10s', type = "Optional")
```


```{r, results = "asis", echo = FALSE}
req_read("Matrix Math", url = 'https://www.mathsisfun.com/algebra/matrix-introduction.html',
            type = "Optional")
```



The main thing to make sure you understand is how **matrix multiplication** (or 
*cross product*) is different from **elementwise multiplication** (or *dot product*).

```{r, error = TRUE}
A <- matrix(c(5, -1, -3, 2), c(2,2))
A

B <- matrix(c(1,2,3,4), c(2,2))
B

A*B
A %*% B
B %*% A

```

In particular, note that **order matters** and that not all matrices can be multiplied
together!

```{r, error = TRUE}
C = matrix(c(5, -1), c(1,2))

A*C
A %*% C
C %*% A

```


### Special Cases

There are a few simple special terms you should know about.

A **square root** of a matrix is the matrix that, if multiplied with itself, 
gets back the original:

$${\bf A}^{1/2} {\bf A}^{1/2} = {\bf A}$$

Note that not every matrix has a valid square root! 

```{r}
sqrt(A)
sqrt(B)
sqrt(C)
```


The **transpose** of a matrix is the "tilted" or "mirror imaged" version, with 
rows and columns swapped:

$$M = \left( \matrix{8 & 3 \\ 4 & 1 \\ 2 & 3} \right)$$

$$M' \text{ or } M^t =  \left( \matrix{8 & 4 & 2 \\ 3 & 1 & 3} \right)$$

```{r}
M <- cbind(c(8, 4, 2), c(3, 1, 3))
M

t(M)
```

There are many situations (especially in statisics) where we want to multiply a
matrix by its transpose.  This is called a **crossproduct**.

```{r}
t(M) %*% M
crossprod(M)


M %*% t(M)
tcrossprod(M)
```

The **inverse** of a matrix is the thing that we can multiply it by to
create the *identity* matrix.


$$ {\bf A}^{-1} {\bf A} = {\bf I}$$

Once again - not all matrices have a valid inverse! In particular, only square
matrices can possibly have inverses.  (Although this is not a guarantee by itself.)

```{r, error = TRUE}
solve(A)

round(solve(A) %*% A)

solve(M)

```


Some matrices (in fact, those with an inverse) have a **determinant**.  This is a
single number that has some mathematical importance to the matrix.  You don't need
to know anything about the underlying math, just how to find it iwth R:

```{r, error = TRUE}
det(A)

det(M)
```


Lastly, the **eigenvalues** and **eigenvectors** of a matrix are certain numbers
and vectors with special properties.  The video in a moment will give you some
intuition behind these.  We won't concern ourselves with the math, but you should
know how to find these with R:

```{r}
eigens <- eigen(A)
eigens

# First "eigenvector" of A
eigens$vectors[,1]

# First "eigenvalue" of A
eigens$values[1]
```

Be careful... you saw this coming... not all matrices have eigenvalues and vectors.

```{r, error = TRUE}
eigen(M)
```



```{r, results = "asis", echo = FALSE}
req_vid("Applications of Matrices", url = 'https://www.youtube.com/embed/rowWM-MijXU')
```

```{r, results = "asis", echo = FALSE}
checkin("Matrix Algebra in R")
```

At [this repository](https://github.com/Cal-Poly-Advanced-R/Matrix-Algebra-Practice), you will find a puzzle using matrix calculations.

The puzzle will ask you to perform a series of steps on a mystery image.  If you
do the steps correctly, the image will be transformed into something recognizeable.

Once you discover the person in the image, you are done!

```{r, results = "asis", echo = FALSE}
canvas_task(quiz_url)
```
