---
title: "Week 5: Webscraping"
format: html
embed-resources: true
editor: visual
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false

source("../scripts/window_dressing.R")

library(tidyverse)
library(flair)
library(lubridate)
library(gganimate)
```

```{r}
#| label: summary-table-of-activities
#| results: asis

make_summary_table(vid_time = "20 min", 
                   reading_time = "30 min", 
                   work_time = "60 min", 
                   canvas_activities = 2)
```



### Web-Scraping

Web pages can be beautiful and rich! HTML provides a wonderful structure with which to house this beauty and richness. Unfortunately, this structure is not the tabular structure that we're used to, but it's still very accessible. In particular, people have developed tools in `R` for working with HTML. One very popular one is the `rvest` package. Take a few minutes to (re-)introduce yourself to web-scraping with `rvest` and working with HTML in `R`.

```{r, results = "asis", echo = FALSE}
req_vid("Introduction to Web-Scraping with rvest", url = 'https://www.youtube.com/watch?v=MFQTHrCiAxA', 
        embed = '<iframe width="400" height="225" style="width: 400px; height: 225px;" src="https://www.youtube.com/embed/MFQTHrCiAxA" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>')
```

I encourage you to following with the video to the point of copying and running the same code that's being demonstrated.

```{r, results = "asis", echo = FALSE}
checkin("rvest")
```

a)  What is the CSS tool called that is used to identify the portions of the webpage we're interested in scraping?

-   CSS Thingy
-   HTML Dissector
-   Selector Gadget
-   HTML2CSS

b)  What types of webpages are scraped in the video? (Make sure you understand the difference between the two, even if you have to look it up)

-   static
-   dynamic

```{r, results = "asis", echo = FALSE}
canvas_task(quiz_url)
```