---
title: "Week 1: Setting up Git and GitHub"
format: 
  html: 
    fontsize: 1em
    mainfont: sans-serif
    number-sections: true
    number-depth: 2
    section-divs: true
    code-block-bg: "#76b5c5"
    highlight-style: monochrome
embed-resources: true
editor: visual
execute: 
  echo: true
  warning: false
  message: false
---

```{r}
#| label: set-up
#| include: false

source("../scripts/window_dressing.R")
```

All your assignments for Stat 541 will be received and submitted via **GitHub**.  This page will take you through some practice activities, so you can get used to everything.


```{r}
#| label: summary-table-of-activities
#| echo: false
#| results: asis

make_summary_table(vid_time = "45 min", 
                   reading_time = "0 min", 
                   work_time = "30-90 min", 
                   canvas_activities = 5)
```

```{r}
#| echo: false
#| results: asis
#| label: extra-resources

extra_rec("Git and GitHub")
```

* [This Video](https://www.youtube.com/embed/w3jLJU7DT5E) provides a cheesy but cute introduction to GitHub.
* [GitHub Desktop](https://desktop.github.com/) is a nice client for interfacing with GitHub.  You do not need it today (or ever), but you may find it makes your life easier.

## Introduction to Git and GitHub

```{r}
#| echo: false
#| results: asis
#| label: required-video-intro-to-git-and-github

req_vid("Intro to Git and GitHub", url = "https://www.youtube.com/embed/BCQHnlnPusY", embed = FALSE)
```

```{r}
#| echo: false
#| results: asis
#| label: check-in-github-account

checkin("Make a GitHub Account")
```

Create an account on [GitHub.com](GitHub.com), if you have not already.

Ideally, your username will be related to your real name, e.g. "atheobold" rather than "SuperKrazy2000".  However, this is not required; choose whatever username you prefer.  In particular, if you want to avoid making your affiliation with this course public in any way, you may want to anonymize your account.

Make sure that your `@calpoly.edu` email is linked to your account.  (This is under the `Email` section of `Settings` on GitHub.com.  You do not need to make your `@calpoly` email your primary email, nor do you need to make it publicly view-able.)

```{r}
#| echo: false
#| results: asis
#| label: check-in-repository

checkin("Make a repository")
```

Create a **private** repository called "About Me".

Create a new text file.  In it, answer the following questions:

- What is your year and major?
- What are you most excited to learn in Stat 541?

Take a screenshot of your repository.  It should look similar to 9:43 of the video.

```{r}
#| echo: false
#| results: asis
#| label: warning-personal-info-in-screenshots

warning_sign("This class will sometimes ask you to take screenshots.  Please make sure that no private information is displayed in the image before you upload it to Canvas!")
```

## Branches and Pull Requests

```{r}
#| echo: false
#| results: asis
#| label: required-video-branching

req_vid("Branching", url = "https://www.youtube.com/embed/oPpnCh7InLY", embed = FALSE)
```

```{r}
#| echo: false
#| results: asis
#| label: check-in-making-a-branch

checkin("Make a branch and pull request")
```

1. Create a new branch of your "About Me" repository.  Name this branch after your favorite animal.

2. In this branch, make a small change to your text file.

3. Create a **Pull Request** to your `master` branch from your new branch.

4. Take a screen shot of your completed pull request.  It should look like 8:14 of the video.

## Forks and Pull Requests

```{r}
#| echo: false
#| results: asis
#| label: required-video-forks-and-pull-requests

req_vid("Forks and pull requests", url = "https://www.youtube.com/embed/_NrSWLQsDL4", embed = FALSE)
```

```{r}
#| echo: false
#| results: asis
#| label: check-in-fork-and-pull

checkin("Fork and pull request someone else's repository")
```

1. Navigate to the [Week 0 Repository](https://github.com/Cal-Poly-Advanced-R/Week_0) for this course.

2. **Fork** the repository into your own account.

3. Edit the poem to add a new line.  The new line should be two sentences or phrases that rhyme. See Dr. Theobold's contribution for an example.

<!-- Update the date for the txt file in the repo! -->

4. Create a **Pull Request** to add your line to the group's poem. Make sure you are pull requesting the original repository owned by the *Cal-Poly-Advanced-R* account, not pull requesting your own forked copy.

5. Once your pull request has been accepted, take a screenshot of that page.  It should look like 11:42 of the video.  

## Issues

```{r}
#| echo: false
#| results: asis
#| label: required-video-issues

req_vid("Issues and Markdown", url = "https://www.youtube.com/embed/WMykv2ZMyEQ", embed = FALSE)
```

```{r}
#| echo: false
#| results: asis
#| label: check-in-make-an-issue

checkin("Comment on an Issue")
```

1. Find the Issue that Dr. Theobold created in the Week 0 Repository.

2. Reply to the question somewhere below. 

3. Use *at least one* Markdown styling trick in your reply.

